---
title: "NPP_QAQC_Daily_Second_Pass"
author: "AJ - KMH"
output: html_document
---

## Sevilleta NPP quality control - Daily data - Second Pass


```{r, echo=FALSE}
# Purpose of this script: This R Markdown document can be used after running the NPP_QAQC_Daily.Rmd report, and manually correcting the daily combined csv file (../output/filled_raw_dataYYYY-MM-DD.csv).  Before running this, make sure that your changes were saved in Excel or wherever you are editing the csv file.

# It will help you to double check that you successfully corrected the data that were reported in the first report.  
```

```{r load libraries, message=F, warning=F, echo=F}
# make sure libraries are loaded for this session
library(ggplot2)
library(lubridate)
library(plyr)
library(reshape2)
library(gridExtra)
library(cowplot)
library(knitr)
library(markdown)
library(xtable)
library(zoo)
library(tidyr)
library(stats)
library(tidyverse)
library(readxl)
```

```{r set file paths, echo=FALSE}
# USER: Where did you write the combined daily output file to?  Enter directory path.
daily.output.file <- "../output/output_daily"

# USER: Where is the large, all-quad data file stored? Enter directory path to file.
all.quad.file <- "../data/data_other/AJ_AllQuads.csv"

# USER: Where is the master species list stored? Enter directory path to files.
master.species.list.file <- "../data/data_other/SevilletaSpeciesList_AJH.csv"
```

```{r read in corrected daily file, echo=FALSE}
filled.data <- read.csv(paste0(daily.output.file, "/filled_raw_data", today(), ".csv", sep = ""), strip.white = T)

# USER: You may need to adjust the data format if it is pulling the data in incorrectly and assigning NA to collection.date.  You can determine the correct format by commenting out the as.Date code below, and rerunning the chunk of code to see how it gets pulled in.  Then, adjust the format = and rerun the chunk
# correct format of date for imported file
filled.data$collection.date <- as.Date(filled.data$collection.date, format = "%m/%d/%y")
```

```{r, echo = F}
# Read in all-quad data file
all.quads <- read.csv(all.quad.file, strip.white = T)

# format
all.quads$kartez <- as.factor(toupper(all.quads$kartez))
all.quads$cover[all.quads$cover < 0] <- NA
all.quads$height[all.quads$height < 0] <- NA
all.quads$date <- as.Date(all.quads$date, format = "%m/%d/%Y")

# Read in a species list file
master.species.list <- read.csv(master.species.list.file, strip.white = T)
```




### Error report - Second Pass

For second pass, check for whether real errors were successfuly manually fixed in the daily data.

The first type of error to consider are possible typos. This code chunk will output a list of possible typos in the site, species, cover, and OTHER columns, by checking these values against a list of "legal" values that we provide.  
Remember to check list of possible "quad codes"  

```{r, echo = F, message = F}
# Create data.frame with all site-treatment-web-plot-subplot-quad-transect combination
all.quad.codes <- join_all(type = "full", dfs = list(
# Fert has plots 1-20, with quads 1-4 for each plot
siteF = expand.grid(site = "F", plot = as.factor(1:20), quad = 1:4),
# Creosote core has webs 1-5, plots N, S, E, W for each web, and quads 1 and 3 for each plot
siteC = expand.grid(site = "C", web = 1:5, plot = c("N", "S", "E", "W"), quad = c(1,3)),
# Black grama core has webs 1 and 5, with plots N, S, E, W, NE for each web, and quads 1 and 3 for each plot
# Black grama core also has web 4, with plots N, S, E, W, NW, NE for each web, and quads 1 and 3 for each plot
siteG = rbind(expand.grid(site = "G", treatment = "C", web = c(1,5), plot = c("N", "S", "E", "W", "NE"), quad = c(1,3)), expand.grid(site = "G", treatment = "C", web = 4, plot = c("NW", "S", "E", "W", "NE"), quad = c(1,3))), 
# Blue grama core has webs 1-5, plots N, S, E, W for each web, and quads 1 and 3 for each plot
siteB = expand.grid(site = "B", web = 1:5, plot = c("N", "E", "W"), quad = c(1,3)),
# burned black grama core has quads 1:15, 26:40 
siteGB = expand.grid(site = "G", treatment = "B", quad = c(1:15,26:40)),
# mixed grass has burned and unburned treatments has quads 1:15, 26:40 
siteMG = expand.grid(site = "MG", treatment = c(NA, "B"), quad = c(1:15,26:40)),
# burned mixed shrub has burned and unburned treatments with quads 1:40 
siteMS = expand.grid(site = "MS", treatment = c(NA, "B"), quad = 1:40),
# monsoon has plots 1-13, with subplots A, B for each plot, and quads 1-2 for each subplot 
siteM = expand.grid(site = "M", plot = as.factor(1:13), subplot = c("A", "B"), quad = 1:2),
# warming has plots 1-40, with quads 1-2 for each plot 
siteW = expand.grid(site = "W", plot = as.factor(1:40), quad = 1:2),
# pinyon juniper core has plots 1-2, with transects V, R, S, N for each plot, and quads 1-5 for each transect 
# pinyon juniper core also has plot 3, with transects A, B, C, D, and quads 1-5 for each transect 
siteP = rbind(expand.grid(site = "P", plot = as.factor(1:2), transect = c("V", "R", "S", "N"), quad = 1:10), expand.grid(site = "P", plot = as.factor(3), transect = c("A", "B", "C", "D"), quad = 1:5))))

all.quad.codes$treatment[is.na(all.quad.codes$treatment)] <- "C"

```

```{r, echo = F}
# Inconsistent collection dates
cat(" \n Inconsistent collection dates: \n")
errors <- unique(filled.data[is.na(filled.data$collection.date) | abs(as.numeric(filled.data$collection.date - mean(filled.data$collection.date, trim=0.1))) > 30, c("field.sheet.name", "site", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


# Rows with missing data
cat(" \n Rows with missing data: \n")
errors <- unique(filled.data[is.na(filled.data$site) |
                               is.na(filled.data$quad) |
                               is.na(filled.data$kartez) |
                               is.na(filled.data$cover) |
                               is.na(filled.data$height) |
                               is.na(filled.data$count) |
                               is.na(filled.data$collection.date) |
                               is.na(filled.data$collector), c("field.sheet.name", "site", "quad", "kartez", "cover", "height", "count", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


```

```{r, echo = F}
# Site-quad errors

# Site names in the filled.data file that don't exist in the all.quads data file
cat(" \n Site names in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!filled.data$site %in% all.quad.codes$site, c("field.sheet.name", "site", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Site-treatment combinations in the filled.data file that don't exist in the all.quads data file
cat(" \n Site-treatment combinations in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!paste(filled.data$site, filled.data$treatment) %in% paste(all.quad.codes$site, all.quad.codes$treatment) ,c("field.sheet.name", "site", "treatment", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Site-treatment-web-plot-subplot-quad-transect-treatment combinations in the filled.data file that don't exist in the all.quads data file
cat(" \n Site-web-plot-subplot-quad-transect combinations in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!paste(filled.data$site, filled.data$treatment, filled.data$web, filled.data$plot, filled.data$subplot, filled.data$transect, filled.data$quad) %in% paste(all.quad.codes$site, all.quad.codes$treatment, all.quad.codes$web, all.quad.codes$plot, all.quad.codes$subplot, all.quad.codes$transect, all.quad.codes$quad), c("field.sheet.name", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "collection.date", "collector")]) 
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


# Missing site-treatment-web-plot-subplot-transect-quad combinations in the filled.data file 
# cat(" \n Missing site-treatment-web-plot-subplot-transect-quad combinations: \n")
# errors <- unique(all.quad.codes[!paste(all.quad.codes$site, all.quad.codes$treatment, all.quad.codes$web, all.quad.codes$plot, all.quad.codes$subplot, all.quad.codes$transect, all.quad.codes$quad) %in% paste(filled.data$site, filled.data$treatment, filled.data$web, filled.data$plot, filled.data$subplot, filled.data$transect, filled.data$quad), c("site", "treatment", "web", "plot", "subplot", "transect", "quad")]) 
# {if (nrow(errors) > 0) kable(errors, format = "pandoc")
# else cat("\n no errors found \n")}


# Duplicate observations of site-web-treatment-plot-subplot-transect-quad combinations in the filled.data file 
# duplicates <- ddply(filled.data, c("site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
#   num.collectors = length(unique(x$collector)),
#   num.dates = length(unique(x$collection.date))))
# duplicates <- duplicates[duplicates$num.collectors > 1 | duplicates$num.dates > 1,]
# cat(" \n Duplicate site-web-treatment-plot-subplot-transect-quad observations: \n")
# errors <- unique(filled.data[paste(filled.data$site, filled.data$treatment, filled.data$web, filled.data$plot, filled.data$subplot, filled.data$transect, filled.data$quad) %in% paste(duplicates$site, duplicates$treatment, duplicates$web, duplicates$plot, duplicates$subplot, duplicates$transect, duplicates$quad), c("field.sheet.name", "collection.date", "collector", "site", "treatment", "web", "plot", "subplot", "transect", "quad")]) 
# {if (nrow(errors) > 0) kable(errors, format = "pandoc")
# else cat("\n no errors found \n")}

# KMH NOTE: Missing combos and duplicates will be added to another script that can be run once data for an entire site has been collected.

```


```{r find typos, echo = F}

# Species names in the filled.data file that don't exist in the all.quads data file
cat(" \n Species names in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!filled.data$kartez %in% all.quads$kartez, c("field.sheet.name", "site", "kartez", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Species names in the filled.data file that don't exist in the species file
cat(" \n Species names in the filled.data file that don't exist in the species file: \n")
errors <- unique(filled.data[!filled.data$kartez %in% master.species.list$kartez, c("field.sheet.name", "site", "kartez", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Cover values that are not in the allowed intervals
cat(" \n Cover values that are not in the allowed intervals: \n")
errors <- unique(filled.data[
  (filled.data$cover > 30 &
     filled.data$kartez %in% filled.data$kartez[filled.data$kartez %in% master.species.list$kartez[master.species.list$family == "Poaceae"]]) |
    (filled.data$cover < 1 & !filled.data$cover %in% c(NA, 0.01, 0.05, 0.1, 0.25, 0.5, 0.75)) | (is.na(filled.data$cover)),
  c("field.sheet.name", "site", "kartez", "cover", "height", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


```







Next, search for entries that are technically legal, but outside the bounds of observations commonly made. For second pass, the user may decide whether or not to have manually corrected these data in the daily file.
Highly improbable:
 -  species never before seen at that site
 -  species cover or height outside of 3 s.d. seen at that site-season
 -  individual measurements add up to a number very different from previous year

```{r quad and site calculations, echo = F, message = F}
# Calculate summary values per quad per species for THIS YEAR
filled.data$count[is.na(filled.data$count)] <- 0
quad.totals <- ddply(filled.data, c("field.sheet.name", "collector", "collection.date", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
  sum.cover = sum(x$cover * x$count, na.rm=T), 
  mean.ind.height = mean(rep(x$height, x$count), na.rm=T)))

# Calculate total cover values per quad per species for ALL PREVIOUS YEARS
allyrs.quad.means <- ddply(all.quads[all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("dataset", "year", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
  allprev.sum.cover = sum(x$cover, na.rm=T),
  allprev.mean.ind.height = mean(x$height, na.rm=T)))
allyrs.quad.means <- ddply(allyrs.quad.means, c("dataset", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
  prev.sum.cover = mean(unique(x$allprev.sum.cover[x$year >= (year(mean(filled.data$collection.date)) - 2)]), na.rm=T),
  prev.mean.height = mean(unique(x$allprev.mean.ind.height[x$year >= (year(mean(filled.data$collection.date)) - 2)]), na.rm=T),
  allprev.sum.cover = mean(x$allprev.sum.cover, na.rm=T),
  allprev.sdsum.cover = sd(x$allprev.sum.cover, na.rm=T),
  allprev.mean.height = mean(x$allprev.mean.ind.height, na.rm=T),
  allprev.sd.height = sd(x$allprev.mean.ind.height, na.rm=T)))

# Calculate average cover and heights per site per species
site.sp.means <- ddply(all.quads[all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("dataset", "site", "treatment", "kartez"), function(x) data.frame(
  sp.mean.cover = mean(x$cover, na.rm=T),
  sp.sd.cover = sd(x$cover, na.rm=T),
  sp.mean.height = mean(x$height, na.rm=T),
  sp.sd.height = sd(x$height, na.rm=T)))

# Calculate total cover values per quad per species for PREVIOUS YEAR
#prevyr.quad.means <- ddply(all.quads[all.quads$year == (year(mean(filled.data$collection.date)) - 1) & all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("dataset", "year", "season", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
#  prev.sum.cover = sum(x$cover, na.rm=T),
#  prev.mean.ind.height = mean(x$height, na.rm=T)))

# Calculate total cover values per quad per species for PREVIOUS TEN YEARS
#allprev.quad.means <- ddply(all.quads[all.quads$year >= (year(mean(filled.data$collection.date)) - 10) & all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
 # allprev.sum.cover = sum(x$cover, na.rm=T) / 10,
#  allprev.mean.ind.height = mean(x$height, na.rm=T),
#  allprev.sd.ind.height = sd(x$cover, na.rm=T)))

# Calculate mean cover values per species per site for THIS YEAR
#site.means <- ddply(quad.totals, c("field.sheet.name", "site", "treatment", "kartez", "collector", "collection.date"), function(x) data.frame(
#  mean.site.sum.cover = sum(x$sum.cover, na.rm=T), 
#  mean.site.cover = mean(x$mean.ind.cover, na.rm=T),
#  mean.site.height = mean(x$mean.ind.height, na.rm=T)))

# Calculate mean cover values per species per site for PREVIOUS TEN YEARS
#prev.site.means <- ddply(all.quads[all.quads$year >= (unique(year(filled.data$collection.date)) - 10) & all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("season", "site", "treatment", "kartez"), function(x) data.frame(
#  prev.mean.site.cover = mean(x$cover, na.rm=T),
#  prev.sd.site.cover = sd(x$cover, na.rm=T),
#  prev.mean.site.height = mean(x$height, na.rm=T),
#  prev.sd.site.height = sd(x$height, na.rm=T)))


quad.compare <- join_all(type = "left", dfs=list(
  quad.totals, allyrs.quad.means, site.sp.means))

```


```{r big changes in quad values, echo = F}
# Cover values that add up to >100% per quad
cat(" \n Cover values that add up to >100% per quad: \n")
errors <- unique(quad.totals[quad.totals$sum.cover > 100, c("field.sheet.name", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez", "sum.cover", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Total cover values that are more than: 
#  2sd from last season's total covers for that quad AND 
#  3sd from the ten-year mean total cover for that quad
cat(" \n Total cover values that changed a lot, per quad: \n")
errors <- unique(quad.compare[
  abs(quad.compare$sum.cover - quad.compare$prev.sum.cover) > quad.compare$allprev.sdsum.cover * 2 & 
    abs(quad.compare$sum.cover - quad.compare$allprev.sum.cover) > quad.compare$allprev.sdsum.cover * 3, 
  c("field.sheet.name", "collector", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez", "sum.cover", "prev.sum.cover", "allprev.sum.cover", "allprev.sdsum.cover")])
errors <- errors %>% filter(!is.na(sum.cover) & !is.na(prev.sum.cover))
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


# Mean height values that are more than: 
#  3sd from last season's mean heights for that quad AND 
#  4sd from the ten-year mean height for that quad
cat(" \n Mean heights that changed a lot, per quad: \n")
errors <- unique(quad.compare[
  abs(quad.compare$mean.ind.height - quad.compare$prev.mean.height) > quad.compare$allprev.sd.height * 3 & 
    abs(quad.compare$mean.ind.height - quad.compare$allprev.mean.height) > quad.compare$allprev.sd.height * 4, 
  c("field.sheet.name", "collector", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez", "mean.ind.height", "prev.mean.height", "allprev.mean.height", "allprev.sd.height")])
errors <- errors %>% filter(!is.na(mean.ind.height) & !is.na(prev.mean.height))
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

#filled.data[filled.data$site == "MG" & filled.data$treatment == "B" & filled.data$quad == 31 & filled.data$kartez == "BOER4",]

#all.quads[all.quads$year == 2016 & all.quads$season == 3 & all.quads$site == "C" & all.quads$treatment == "C" & all.quads$web == 5 & all.quads$plot == "S" & all.quads$kartez == "LATR2",]


```

After you are finished running the second pass, you may either need to make more manual modifications to the daily file and rerun this report, or you are done for the day.

Remember to move today's daily files to the data_daily_archive directories.  If you want to save a copy of the daily report or the second pass report, open it/them in a web browser, right click on the body, select 'Save Page as...', add today's date to the file name, and save a copy to the project's output/report_archive directory.






