---
title: "NPP_QAQC_Daily"
author: "AJ - KMH"
output:
  html_document: default
  pdf_document: default
---

## Sevilleta NPP quality control - Daily data
 
Date of report:
```{r, echo=FALSE, warning=FALSE}
# REPORT DATE:
Sys.time()
```

```{r, echo = FALSE}
# Purpose of this script: This script reads in daily raw excel files from NPP quad observations in the field and outputs one filled, merged, and amended daily file, as well as a report on possible errors within that daily file. In addition to the raw observation files, it utilizes a file of all previous NPP observations made at the Sevilleta.  
```

```{r load libraries convert files specify directories, message=F, warning=F, echo=F}
# !!!INITIAL STEPS FOR USER: Make sure libraries are up to date, and that directory paths are correctly set.  I place "USER" in locations where the user may need to update information.

# This initial code will not appear in the report output.  

# USER: Update libary packages as needed.
# Load R packages that will be used
# This list is long because Alesia is lazy 
library(ggplot2)
library(lubridate)
library(plyr)
library(reshape2)
library(gridExtra)
library(cowplot)
library(knitr)
library(markdown)
library(xtable)
library(zoo)
library(tidyr)
library(stats)
library(tidyverse)
library(readxl)


# Convert the daily .xlsx files to .csv files - you may need to change the directory paths
# USER: Where are the daily xlsx files are located? Enter directory path to files.
location_xlsx.data.sheets <- 
  list.files("C:/Users/sevfi/Desktop/NPP_QAQC/NPP_QAQC/data/data_daily/xlsx_format")

for(i in location_xlsx.data.sheets) {
  file.list <- lapply(location_xlsx.data.sheets, function(x) gsub(".xlsx", "", x))
}

for(i in file.list){
filepath <- file.path("C:/Users/sevfi/Desktop/NPP_QAQC/NPP_QAQC/data/data_daily/xlsx_format",paste(i,".xlsx",sep=""))
assign(i, read_xlsx(filepath))
}

# USER: Enter the directory path where the daily csv formatted files will be written.
for(i in ls(pattern = "npp")) {
   write.csv(get(i), paste("C:/Users/sevfi/Desktop/NPP_QAQC/NPP_QAQC/data/data_daily/csv_format/", i, ".csv", sep = ""), row.names = FALSE)
}

rm(list=ls(all=TRUE))



# USER: Where are the daily raw csv data files stored? Enter directory path to files.
location_raw.data.sheets <- "C:/Users/sevfi/Desktop/NPP_QAQC/NPP_QAQC/data/data_daily/csv_format"

# USER: Where is the large, all-quad data file stored? Enter directory path to file.
all.quad.file <- "C:/Users/sevfi/Desktop/NPP_QAQC/NPP_QAQC/data/data_other/AJ_AllQuads.csv"

# USER: Where is the master species list stored? Enter directory path to files.
master.species.list.file <- "C:/Users/sevfi/Desktop/NPP_QAQC/NPP_QAQC/data/data_other/SevilletaSpeciesList_AJH.csv"

# USER: Where will you write the combined daily output file to?  Enter directory path.
daily.output.file <- "C:/Users/sevfi/Desktop/NPP_QAQC/NPP_QAQC/output/output_daily/"

```




```{r read in raw files, echo = F}
# An important step is to read in all of the raw daily csv files, merge them together, and fill in blank site, plot, or quad columns
# In the field, data collectors don't manually fill in every row in the file, because this is a waste of time. This loop reads through each row of the data file and, if cells within the site, plot, or quad columns are empty, fills that cell with information from the previous cell. 


# Read in raw data files (CSV format)
raw.data.sheets <- list.files(location_raw.data.sheets, pattern = glob2rx("npp*csv"), full.names = T)


# Define legal column names
legal.col.names <- c("site", "treatment", "quad", "kartez", "obs", "cover", "height", "count", "comments", "field.sheet.name", "collection.date", "collector", "web", "plot", "subplot", "transect", "field_comments", "qaqc_comments")

for (i in 1:length(raw.data.sheets)) {
  raw.data.file <- read.csv(raw.data.sheets[i], header = T, strip.white = T, na.strings = c("NA", ""))
  # Delete unnamed columns
  if ("x" %in% colnames(raw.data.file)) {
    raw.data.file <- subset(raw.data.file, select = -x) }
  if ("X" %in% colnames(raw.data.file)) {
    raw.data.file <- subset(raw.data.file, select = -X) }
  
  # extract file name info
  raw.data.file$field.sheet.name <- as.factor(list.files(location_raw.data.sheets, pattern=glob2rx("npp*csv"))[i])
  field.file.name <- strsplit(list.files(location_raw.data.sheets, pattern = glob2rx("npp*csv"))[i], split = "[.]")[[1]]
  
  # make column names lowercase and check if they are legal
  colnames(raw.data.file) <- tolower(colnames(raw.data.file))
  colnames(raw.data.file)[colnames(raw.data.file) == "trt"] <- "treatment"
  colnames(raw.data.file)[colnames(raw.data.file) == "spp"] <- "kartez"
  colnames(raw.data.file)[colnames(raw.data.file) == "species"] <- "kartez"
  colnames(raw.data.file)[colnames(raw.data.file) == "comm"] <- "comments"
  colnames(raw.data.file)[colnames(raw.data.file) == "trans"] <- "transect"

  # Break loop if any "illegal" column names exist
  if (any(!colnames(raw.data.file) %in% legal.col.names)) {
    print(paste("A column name in file", raw.data.file$field.sheet.name[1], "needs to be corrected before continuing."))
    break
  }
  
  # Extract collection date from file name
  raw.data.file$collection.date <- as.Date(paste(
    field.file.name[2], # month
    field.file.name[3], # day of month
    substr(field.file.name[4], nchar(field.file.name[4])-1, nchar(field.file.name[4]))), # last two digits of year
    format = "%m %d %y")
  
  # Extract three-letter collector initials from file name
  raw.data.file$collector <- as.factor(substr(field.file.name[5],1,3))

  # Fill in blank site, plot, and quad cells
for (j in 1:nrow(raw.data.file)) {
  if("site" %in% colnames(raw.data.file)) if (is.na(raw.data.file$site[j])) {
    raw.data.file$site[j] <- raw.data.file$site[j-1]}
  if("web" %in% colnames(raw.data.file)) if (is.na(raw.data.file$web[j])) {
    raw.data.file$web[j] <- raw.data.file$web[j-1]}
  if("plot" %in% colnames(raw.data.file)) if (is.na(raw.data.file$plot[j])) {
    raw.data.file$plot[j] <- raw.data.file$plot[j-1]}
  if("subplot" %in% colnames(raw.data.file)) if (is.na(raw.data.file$subplot[j])) {
    raw.data.file$subplot[j] <- raw.data.file$subplot[j-1]}
  if("transect" %in% colnames(raw.data.file)) if (is.na(raw.data.file$transect[j])) {
    raw.data.file$transect[j] <- raw.data.file$transect[j-1]}
  if("treatment" %in% colnames(raw.data.file)) if (is.na(raw.data.file$treatment[j])) {
    raw.data.file$treatment[j] <- raw.data.file$treatment[j-1]}
  if("quad" %in% colnames(raw.data.file)) if (is.na(raw.data.file$quad[j])) {
    raw.data.file$quad[j] <- raw.data.file$quad[j-1]}
  if("kartez" %in% colnames(raw.data.file)) if (is.na(raw.data.file$kartez[j])) {
    raw.data.file$kartez[j] <- raw.data.file$kartez[j-1]}
}
  
  # Merge all data files together
  if (i == 1) raw.data <- raw.data.file
  else raw.data <- join(raw.data, raw.data.file, by=intersect(colnames(raw.data), colnames(raw.data.file)), type="full")
}


# Add variables if they are not present in the data
if ("web" %in% colnames(raw.data)) {
  raw.data <- raw.data
} else {
  raw.data$web <- NA
}

if ("plot" %in% colnames(raw.data)) {
  raw.data <- raw.data
} else {
  raw.data$plot <- NA
}

if ("subplot" %in% colnames(raw.data)) {
  raw.data <- raw.data
} else {
  raw.data$subplot <- NA
}

if ("transect" %in% colnames(raw.data)) {
  raw.data <- raw.data
} else {
  raw.data$transect <- NA
}

if ("treatment" %in% colnames(raw.data)) {
  raw.data <- raw.data
} else {
  raw.data$treatment <- NA
}

rm(raw.data.file)
```



```{r, echo = F}
# Read in all-quad data file
all.quads <- read.csv(all.quad.file, strip.white = T)

# format
all.quads$kartez <- as.factor(toupper(all.quads$kartez))
all.quads$cover[all.quads$cover < 0] <- NA
all.quads$height[all.quads$height < 0] <- NA
all.quads$date <- as.Date(all.quads$date, format = "%m/%d/%Y")

# Read in a species list file
master.species.list <- read.csv(master.species.list.file, strip.white = T)
```


```{r quick fixes, echo = F}
# There are some "errors" that can be fixed easily and without any special consideration.  

# Capitalize site and species names
# raw.data$site <- as.factor(toupper(raw.data$site))
# raw.data$kartez <- as.factor(toupper(raw.data$kartez))
# raw.data$treatment <- as.factor(toupper(raw.data$treatment))
# raw.data$plot <- as.factor(toupper(raw.data$plot))
# raw.data$subplot <- as.factor(toupper(raw.data$subplot))
# raw.data$transect <- as.factor(toupper(raw.data$transect))

# Capitalize site and species names
# not all variable may be present in all daily data sets, but this will change them to upper if present
if("site" %in% colnames(raw.data)) {raw.data$site <- as.factor(toupper(raw.data$site))}
if("kartez" %in% colnames(raw.data)) {raw.data$kartez <- as.factor(toupper(raw.data$kartez))}
if("treatment" %in% colnames(raw.data)) {raw.data$treatment <- as.factor(toupper(raw.data$treatment))}
if("plot" %in% colnames(raw.data)) {raw.data$plot<- as.factor(toupper(raw.data$plot))}
if("subplot" %in% colnames(raw.data)) {raw.data$subplot <- as.factor(toupper(raw.data$subplot))}
if("transect" %in% colnames(raw.data)) {raw.data$transect <- as.factor(toupper(raw.data$transect))}


# Fix some common site name mistakes
raw.data$site <- as.factor(mapvalues(
  as.character(raw.data$site), 
  from = "FALSE", to = "F", warn_missing = F))
raw.data$site <- as.factor(mapvalues(
  as.character(raw.data$site), 
  from = "FERT", to = "F", warn_missing = F))
raw.data$site <- as.factor(mapvalues(
  as.character(raw.data$site), 
  from = "WARM", to = "W", warn_missing = F))
raw.data$site <- as.factor(mapvalues(
  as.character(raw.data$site), 
  from = "WARMING", to = "W", warn_missing = F))
raw.data$site <- as.factor(mapvalues(
  as.character(raw.data$site), 
  from = "PJ", to = "P", warn_missing = F))

# Change treatment NA or "U" (unburned) to C (control)
raw.data$treatment <- as.factor(mapvalues(
  as.character(raw.data$treatment), 
  from = c("U", NA), to = c("C", "C"), warn_missing = F))

# Delete commas in comment column
raw.data$field_comments <- gsub(x = raw.data$field_comments,
                          pattern = ",",
                          replacement = ".")


```

Initial summary and look at the values of the variables, if present in the data. Some files will not contain all variables that are possible.

Summary:
```{r, echo=FALSE}
summary(raw.data)
```
site:
```{r, echo=FALSE}
# site
table(as.factor(raw.data$site))
```
treatment:
```{r, echo=FALSE}
# treatment
table(as.factor(raw.data$treatment))
```
quad:
```{r, echo=FALSE}
# quad
table(as.factor(raw.data$quad))
```
kartez:
```{r, echo=FALSE}
# kartez
table(as.factor(raw.data$kartez))
```
obs:
```{r, echo=FALSE}
# obs
table(as.factor(raw.data$obs))
```
cover:
```{r, echo=FALSE}
# cover
table(as.factor(raw.data$cover))
```
height:
```{r, echo=FALSE}
# height
table(as.factor(raw.data$height))
```
count:
```{r, echo=FALSE}
# count
table(as.factor(raw.data$count))
```
collection.date:
```{r, echo=FALSE}
# collection.date
file.dates <- unique(raw.data$collection.date)
if (file.dates != today()) {print(paste("HEY!: There are dates in the daily files that are not today's date"))}

table(as.factor(raw.data$collection.date))
```
collector:
```{r, echo=FALSE}
# collector
table(as.factor(raw.data$collector))
```
web:
```{r, echo=FALSE}
# web
table(as.factor(raw.data$web))
```
plot:
```{r, echo=FALSE}
# plot
table(as.factor(raw.data$plot))
```
subplot:
```{r, echo=FALSE}
# subplot
table(as.factor(raw.data$subplot))
```
transect:
```{r, echo=FALSE}
# transect
table(as.factor(raw.data$transect))
```


```{r write out initial raw file, echo = F}
# Write out this new data file. 

# Write out data file
# write.out.file <- paste0(location_raw.data.sheets, "filled_raw_data", months(mean(raw.data$collection.date), abb=T), year(mean(raw.data$collection.date)), ".csv")
write.out.file <- paste0(daily.output.file, "/filled_raw_data", today(), ".csv", sep = "")

# print(write.out.file)
write.csv(raw.data, write.out.file, quote = F, row.names = F)

# USER: Will need to change to the correct directory path
 # filled.data <- read.csv("C:/Users/sevuser/Desktop/UNM/2017 Data/NPP/after R/filled_raw_dataAug2017.csv", strip.white = T)
filled.data <- read.csv(paste0(daily.output.file, "filled_raw_data", today(), ".csv", sep = ""), strip.white = T)

#filled.data$collection.date <- as.Date(filled.data$collection.date, format = "%m/%d/%Y")
# kmh change - filled.data reads back into R in %Y-%m-%d format
filled.data$collection.date <- as.Date(filled.data$collection.date, format = "%Y-%m-%d")


#filled.data <- read.csv(write.out.file, strip.white = T)

#filled.data$collection.date <- as.Date(filled.data$collection.date)
```



### Error report

Now we can search for real errors. These errors are not fixed within this script, but simply reported. They should be evaluated by hand and corrected in the raw data or the new file, if necessary.
The first type of error to consider are possible typos. This code chunk will output a list of possible typos in the site, species, cover, and OTHER columns, by checking these values against a list of "legal" values that we provide.  
Remember to check list of possible "quad codes"  

```{r, echo = F, message = F}
# Create data.frame with all site-treatment-web-plot-subplot-quad-transect combination
all.quad.codes <- join_all(type = "full", dfs = list(
# Fert has plots 1-20, with quads 1-4 for each plot
siteF = expand.grid(site = "F", plot = as.factor(1:20), quad = 1:4),
# Creosote core has webs 1-5, plots N, S, E, W for each web, and quads 1 and 3 for each plot
siteC = expand.grid(site = "C", web = 1:5, plot = c("N", "S", "E", "W"), quad = c(1,3)),
# Black grama core has webs 1 and 5, with plots N, S, E, W, NE for each web, and quads 1 and 3 for each plot
# Black grama core also has web 4, with plots N, S, E, W, NW, NE for each web, and quads 1 and 3 for each plot
siteG = rbind(expand.grid(site = "G", treatment = "C", web = c(1,5), plot = c("N", "S", "E", "W", "NE"), quad = c(1,3)), expand.grid(site = "G", treatment = "C", web = 4, plot = c("NW", "S", "E", "W", "NE"), quad = c(1,3))), 
# Blue grama core has webs 1-5, plots N, S, E, W for each web, and quads 1 and 3 for each plot
siteB = expand.grid(site = "B", web = 1:5, plot = c("N", "E", "W"), quad = c(1,3)),
# burned black grama core has quads 1:15, 26:40 
siteGB = expand.grid(site = "G", treatment = "B", quad = c(1:15,26:40)),
# mixed grass has burned and unburned treatments has quads 1:15, 26:40 
siteMG = expand.grid(site = "MG", treatment = c(NA, "B"), quad = c(1:15,26:40)),
# burned mixed shrub has burned and unburned treatments with quads 1:40 
siteMS = expand.grid(site = "MS", treatment = c(NA, "B"), quad = 1:40),
# monsoon has plots 1-13, with subplots A, B for each plot, and quads 1-2 for each subplot 
siteM = expand.grid(site = "M", plot = as.factor(1:13), subplot = c("A", "B"), quad = 1:2),
# warming has plots 1-40, with quads 1-2 for each plot 
siteW = expand.grid(site = "W", plot = as.factor(1:40), quad = 1:2),
# pinyon juniper core has plots 1-2, with transects V, R, S, N for each plot, and quads 1-5 for each transect 
# pinyon juniper core also has plot 3, with transects A, B, C, D, and quads 1-5 for each transect 
siteP = rbind(expand.grid(site = "P", plot = as.factor(1:2), transect = c("V", "R", "S", "N"), quad = 1:10), expand.grid(site = "P", plot = as.factor(3), transect = c("A", "B", "C", "D"), quad = 1:5))))

all.quad.codes$treatment[is.na(all.quad.codes$treatment)] <- "C"

```

```{r, echo = F}
# Inconsistent collection dates
cat(" \n Inconsistent collection dates: \n")
errors <- unique(filled.data[is.na(filled.data$collection.date) | abs(as.numeric(filled.data$collection.date - mean(filled.data$collection.date, trim=0.1))) > 30, c("field.sheet.name", "site", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


# Rows with missing data
cat(" \n Rows with missing data: \n")
errors <- unique(filled.data[is.na(filled.data$site) |
                               is.na(filled.data$quad) |
                               is.na(filled.data$kartez) |
                               is.na(filled.data$cover) |
                               is.na(filled.data$height) |
                               is.na(filled.data$count) |
                               is.na(filled.data$collection.date) |
                               is.na(filled.data$collector), c("field.sheet.name", "site", "quad", "kartez", "cover", "height", "count", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


```

```{r, echo = F}
# Site-quad errors

# Site names in the filled.data file that don't exist in the all.quads data file
cat(" \n Site names in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!filled.data$site %in% all.quad.codes$site, c("field.sheet.name", "site", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Site-treatment combinations in the filled.data file that don't exist in the all.quads data file
cat(" \n Site-treatment combinations in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!paste(filled.data$site, filled.data$treatment) %in% paste(all.quad.codes$site, all.quad.codes$treatment) ,c("field.sheet.name", "site", "treatment", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Site-treatment-web-plot-subplot-quad-transect-treatment combinations in the filled.data file that don't exist in the all.quads data file
cat(" \n Site-web-plot-subplot-quad-transect combinations in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!paste(filled.data$site, filled.data$treatment, filled.data$web, filled.data$plot, filled.data$subplot, filled.data$transect, filled.data$quad) %in% paste(all.quad.codes$site, all.quad.codes$treatment, all.quad.codes$web, all.quad.codes$plot, all.quad.codes$subplot, all.quad.codes$transect, all.quad.codes$quad), c("field.sheet.name", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "collection.date", "collector")]) 
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


# Missing site-treatment-web-plot-subplot-transect-quad combinations in the filled.data file 
# cat(" \n Missing site-treatment-web-plot-subplot-transect-quad combinations: \n")
# errors <- unique(all.quad.codes[!paste(all.quad.codes$site, all.quad.codes$treatment, all.quad.codes$web, all.quad.codes$plot, all.quad.codes$subplot, all.quad.codes$transect, all.quad.codes$quad) %in% paste(filled.data$site, filled.data$treatment, filled.data$web, filled.data$plot, filled.data$subplot, filled.data$transect, filled.data$quad), c("site", "treatment", "web", "plot", "subplot", "transect", "quad")]) 
# {if (nrow(errors) > 0) kable(errors, format = "pandoc")
# else cat("\n no errors found \n")}


# Duplicate observations of site-web-treatment-plot-subplot-transect-quad combinations in the filled.data file 
# duplicates <- ddply(filled.data, c("site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
#   num.collectors = length(unique(x$collector)),
#   num.dates = length(unique(x$collection.date))))
# duplicates <- duplicates[duplicates$num.collectors > 1 | duplicates$num.dates > 1,]
# cat(" \n Duplicate site-web-treatment-plot-subplot-transect-quad observations: \n")
# errors <- unique(filled.data[paste(filled.data$site, filled.data$treatment, filled.data$web, filled.data$plot, filled.data$subplot, filled.data$transect, filled.data$quad) %in% paste(duplicates$site, duplicates$treatment, duplicates$web, duplicates$plot, duplicates$subplot, duplicates$transect, duplicates$quad), c("field.sheet.name", "collection.date", "collector", "site", "treatment", "web", "plot", "subplot", "transect", "quad")]) 
# {if (nrow(errors) > 0) kable(errors, format = "pandoc")
# else cat("\n no errors found \n")}

# KMH NOTE: Missing combos and duplicates will be added to another script that can be run once data for an entire site has been collected.

```


```{r find typos, echo = F}

# Species names in the filled.data file that don't exist in the all.quads data file
cat(" \n Species names in the filled.data file that don't exist in the all.quads data file: \n")
errors <- unique(filled.data[!filled.data$kartez %in% all.quads$kartez, c("field.sheet.name", "site", "kartez", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Species names in the filled.data file that don't exist in the species file
cat(" \n Species names in the filled.data file that don't exist in the species file: \n")
errors <- unique(filled.data[!filled.data$kartez %in% master.species.list$kartez, c("field.sheet.name", "site", "kartez", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Cover values that are not in the allowed intervals
cat(" \n Cover values that are not in the allowed intervals: \n")
errors <- unique(filled.data[
  (filled.data$cover > 30 &
     filled.data$kartez %in% filled.data$kartez[filled.data$kartez %in% master.species.list$kartez[master.species.list$family == "Poaceae"]]) |
    (filled.data$cover < 1 & !filled.data$cover %in% c(NA, 0.01, 0.05, 0.1, 0.25, 0.5, 0.75)) | (is.na(filled.data$cover)),
  c("field.sheet.name", "site", "kartez", "cover", "height", "collection.date", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


```







Next, we can search for entries that are technically legal, but outside the bounds of observations commonly made. 
Highly improbable:
 -  species never before seen at that site
 -  species cover or height outside of 3 s.d. seen at that site-season
 -  individual measurements add up to a number very different from previous year

```{r quad and site calculations, echo = F, message = F}
# Calculate summary values per quad per species for THIS YEAR
filled.data$count[is.na(filled.data$count)] <- 0
quad.totals <- ddply(filled.data, c("field.sheet.name", "collector", "collection.date", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
  sum.cover = sum(x$cover * x$count, na.rm=T), 
  mean.ind.height = mean(rep(x$height, x$count), na.rm=T)))

# Calculate total cover values per quad per species for ALL PREVIOUS YEARS
allyrs.quad.means <- ddply(all.quads[all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("dataset", "year", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
  allprev.sum.cover = sum(x$cover, na.rm=T),
  allprev.mean.ind.height = mean(x$height, na.rm=T)))
allyrs.quad.means <- ddply(allyrs.quad.means, c("dataset", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
  prev.sum.cover = mean(unique(x$allprev.sum.cover[x$year >= (year(mean(filled.data$collection.date)) - 2)]), na.rm=T),
  prev.mean.height = mean(unique(x$allprev.mean.ind.height[x$year >= (year(mean(filled.data$collection.date)) - 2)]), na.rm=T),
  allprev.sum.cover = mean(x$allprev.sum.cover, na.rm=T),
  allprev.sdsum.cover = sd(x$allprev.sum.cover, na.rm=T),
  allprev.mean.height = mean(x$allprev.mean.ind.height, na.rm=T),
  allprev.sd.height = sd(x$allprev.mean.ind.height, na.rm=T)))

# Calculate average cover and heights per site per species
site.sp.means <- ddply(all.quads[all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("dataset", "site", "treatment", "kartez"), function(x) data.frame(
  sp.mean.cover = mean(x$cover, na.rm=T),
  sp.sd.cover = sd(x$cover, na.rm=T),
  sp.mean.height = mean(x$height, na.rm=T),
  sp.sd.height = sd(x$height, na.rm=T)))

# Calculate total cover values per quad per species for PREVIOUS YEAR
#prevyr.quad.means <- ddply(all.quads[all.quads$year == (year(mean(filled.data$collection.date)) - 1) & all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("dataset", "year", "season", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
#  prev.sum.cover = sum(x$cover, na.rm=T),
#  prev.mean.ind.height = mean(x$height, na.rm=T)))

# Calculate total cover values per quad per species for PREVIOUS TEN YEARS
#allprev.quad.means <- ddply(all.quads[all.quads$year >= (year(mean(filled.data$collection.date)) - 10) & all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez"), function(x) data.frame(
 # allprev.sum.cover = sum(x$cover, na.rm=T) / 10,
#  allprev.mean.ind.height = mean(x$height, na.rm=T),
#  allprev.sd.ind.height = sd(x$cover, na.rm=T)))

# Calculate mean cover values per species per site for THIS YEAR
#site.means <- ddply(quad.totals, c("field.sheet.name", "site", "treatment", "kartez", "collector", "collection.date"), function(x) data.frame(
#  mean.site.sum.cover = sum(x$sum.cover, na.rm=T), 
#  mean.site.cover = mean(x$mean.ind.cover, na.rm=T),
#  mean.site.height = mean(x$mean.ind.height, na.rm=T)))

# Calculate mean cover values per species per site for PREVIOUS TEN YEARS
#prev.site.means <- ddply(all.quads[all.quads$year >= (unique(year(filled.data$collection.date)) - 10) & all.quads$season == (unique(semester(filled.data$collection.date)) + 1),], c("season", "site", "treatment", "kartez"), function(x) data.frame(
#  prev.mean.site.cover = mean(x$cover, na.rm=T),
#  prev.sd.site.cover = sd(x$cover, na.rm=T),
#  prev.mean.site.height = mean(x$height, na.rm=T),
#  prev.sd.site.height = sd(x$height, na.rm=T)))


quad.compare <- join_all(type = "left", dfs=list(
  quad.totals, allyrs.quad.means, site.sp.means))

```


```{r big changes in quad values, echo = F}
# Cover values that add up to >100% per quad
cat(" \n Cover values that add up to >100% per quad: \n")
errors <- unique(quad.totals[quad.totals$sum.cover > 100, c("field.sheet.name", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez", "sum.cover", "collector")])
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

# Total cover values that are more than: 
#  2sd from last season's total covers for that quad AND 
#  3sd from the ten-year mean total cover for that quad
cat(" \n Total cover values that changed a lot, per quad: \n")
errors <- unique(quad.compare[
  abs(quad.compare$sum.cover - quad.compare$prev.sum.cover) > quad.compare$allprev.sdsum.cover * 2 & 
    abs(quad.compare$sum.cover - quad.compare$allprev.sum.cover) > quad.compare$allprev.sdsum.cover * 3, 
  c("field.sheet.name", "collector", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez", "sum.cover", "prev.sum.cover", "allprev.sum.cover", "allprev.sdsum.cover")])
errors <- errors %>% filter(!is.na(sum.cover) & !is.na(prev.sum.cover))
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}


# Mean height values that are more than: 
#  3sd from last season's mean heights for that quad AND 
#  4sd from the ten-year mean height for that quad
cat(" \n Mean heights that changed a lot, per quad: \n")
errors <- unique(quad.compare[
  abs(quad.compare$mean.ind.height - quad.compare$prev.mean.height) > quad.compare$allprev.sd.height * 3 & 
    abs(quad.compare$mean.ind.height - quad.compare$allprev.mean.height) > quad.compare$allprev.sd.height * 4, 
  c("field.sheet.name", "collector", "site", "treatment", "web", "plot", "subplot", "transect", "quad", "kartez", "mean.ind.height", "prev.mean.height", "allprev.mean.height", "allprev.sd.height")])
errors <- errors %>% filter(!is.na(mean.ind.height) & !is.na(prev.mean.height))
{if (nrow(errors) > 0) kable(errors, format = "pandoc")
else cat("\n no errors found \n")}

#filled.data[filled.data$site == "MG" & filled.data$treatment == "B" & filled.data$quad == 31 & filled.data$kartez == "BOER4",]

#all.quads[all.quads$year == 2016 & all.quads$season == 3 & all.quads$site == "C" & all.quads$treatment == "C" & all.quads$web == 5 & all.quads$plot == "S" & all.quads$kartez == "LATR2",]


```

After you are finished running the daily report and second pass, remember to move today's daily files to the data_daily_archive directories.  If you want to save a copy of the daily report, open it in a web browser, right click on the body, select 'Save Page as...', add today's date to the file name, and save a copy to the project's output/report_archive directory.
